---
title: "Api Regression"
author: "Karla Olivier"
output: html_document
---

## 0. Data and package Import

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

```{r packages}
library(readr)
library(stargazer)
library(lmtest)
library(plm)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(readxl)
```

```{r}

## import of the data from python
data <- read.csv("df_data.csv") |>
  mutate(
    date = mdy(date),
    year = year(date)
  ) |> 
  filter(year > 2006)

## import of the forcast data downloaded from webpage
df_forecast <- read_excel("chart-gallery.xlsx", sheet = "1") |> 
  dplyr::select(c("...2", "...4")) |>  ## selecting the columnn i want to stay with
  dplyr::rename(date = "...2", forecast = "...4") |> ##renaming
  filter(!is.na(date) & !is.na(forecast)) |> ## deleting the NANs 
  filter(date != "Month") |>  #deleting the word month that appears for some reason          
  mutate(date = as.Date(as.numeric(date), origin = "1899-12-30")) ## fixing the date format 

df_final <- data  ## i created a new variable not to damage the original one

df_final <- bind_rows( # binding the forcast with the df by rows
  df_final,
  df_forecast |> # inside that dataframe i look for the last observation to see where to start
    filter(date > last(df_final$date))|> 
    mutate(wti = as.numeric(forecast))) |> # create a column an numeric called wti
    dplyr::select(-forecast) |> # i delete this because now i have my wti columns
  arrange(date) |> 
  fill(cpi, core_cpi, diesel, .direction = "down") |> # fill with the last obs of each variables to the rest of the 2026
  mutate(
    year           = year(date), #create new variables 
    wti_gallon     = wti / 42, # covert per barrel to per gallon
    spread_wti_diesel = diesel - wti_gallon) |>
  filter(year < 2027)
```

```{r}
library(readxl)

yoy_data <- df_final |> ## this chunk transform my data into a yoy change and add a covid dummy
  arrange(date) |>
  mutate(
    across(
      c(cpi, core_cpi, diesel, wti, wti_gallon, spread_wti_diesel),
      ~ 100 * (log(.) - log(dplyr::lag(., 12)))),
    covid = as.integer(year >= 2020 & year <= 2022)) |>
  filter(year > 2007)

```

## 1. Data graphs

```{r}
source("theme_bcrd.R")
ggplot(yoy_data, aes(x = date, y = wti)) +
  geom_line(linewidth = 0.8) +
  annotate("rect",
    xmin = as.Date("2026-03-01"), xmax = max(yoy_data$date),
    ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.4) + ## shaded gray area
  geom_hline(yintercept = mean(yoy_data$wti, na.rm = TRUE),
    linetype = "dashed", color = "darkred") +
  labs(title = "WTI YoY (%)", x = "years", y = "%") +
  theme_bcrd +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_text(family = "Gotham Book", size = 10))

```

```{r}

df_plot <- df_final |>
  filter(year(date) < 2026)

ggplot(df_plot, aes(x = date, y = spread_wti_diesel)) +
  geom_line(linewidth = 0.8) +
  geom_hline(
    yintercept = mean(df_plot$spread_wti_diesel, na.rm = TRUE),
    linetype = "dashed",
    color = "darkred"
  ) +
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  labs(
    title = "Spread Diesel and WTI per Gallon",
    x = "Year",
    y = "Per Gallon"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## 2. OLS Regressions

```{r}

yoy_data_r <- yoy_data |>
  filter(year(date) < 2026)

model1 <- lm(cpi ~ diesel + spread_wti_diesel + covid, data = yoy_data_r)
model2 <- lm(cpi ~ wti    + spread_wti_diesel + covid, data = yoy_data_r)
model3 <- lm(core_cpi ~ diesel + spread_wti_diesel + covid, data = yoy_data_r)
model4 <- lm(core_cpi ~ wti    + spread_wti_diesel + covid, data = yoy_data_r)

# summary(model1)
# summary(model2)
# summary(model3)
# summary(model4)
```

## 3. Yearly OLS Regressions

```{r}

results <- data.frame()

for (y in sort(unique(yoy_data$year))) {  ## get each unique year from the data 
  temp <- subset(yoy_data, year == y) #for each year it subset the data 
    for (dep in c("core_cpi", "cpi")) { 
      for (ind in c("diesel", "wti")) {
        fit        <- lm(as.formula(paste(dep, "~", ind, "+ covid")), data = temp) ## all posible combinations given the dep and ind variables
        ct         <- broom::tidy(fit) # helps mme create a nice df for each output 
        ct$year    <- y; ct$dependent <- dep; ct$independent <- ind # this columns help me se which model it is 
        results    <- rbind(results, ct) # adding the ct in the results df. 
      }
    }
  }

## here i use the function for each model
diesel_core     <- subset(results, dependent == "core_cpi" & independent == "diesel" & term == "diesel") |> filter(year < 2026)
diesel_headline <- subset(results, dependent == "cpi"      & independent == "diesel" & term == "diesel") |> filter(year < 2026)
wti_core        <- subset(results, dependent == "core_cpi" & independent == "wti"    & term == "wti")
wti_headline    <- subset(results, dependent == "cpi"      & independent == "wti"    & term == "wti")

```

```{r}
source("theme_bcrd.R")
## i created a fucntion to create graphs
plot_coef <- function(df, title) {
  ggplot(df, aes(x = as.numeric(year), y = estimate)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "darkred") + # a line in the 0 
    scale_x_continuous(breaks = seq(min(df$year), max(df$year), by = 1)) +
    labs(title = title, x = "Year", y = "Coefficient") +
    theme_bcrd +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_text(family = "Gotham Book", size = 10))
}

plot_coef(diesel_core,     "Diesel Impact on Core Inflation (Yearly)")
plot_coef(diesel_headline, "Diesel Impact on Headline Inflation (Yearly)")
plot_coef(wti_core,        "WTI Impact on Core Inflation (Yearly)")
plot_coef(wti_headline,    "WTI Impact on Headline Inflation (Yearly)")


```

## 4. Stargazer with Robust Standard Error

The strongest passthrough comes in the headline inflation that receives a 0.02pp WTI and 0.03pp diesel, while the impact in the core inflation is less. The diesel–WTI spread is also significant in most specifications, suggesting additional explanatory power for inflation dynamics beyond individual price levels.

```{r, results='asis'}
ols_r1 <- coeftest(model1, vcov = vcovHC(model1, type = "HC0"))
ols_r2 <- coeftest(model2, vcov = vcovHC(model2, type = "HC0"))
ols_r3 <- coeftest(model3, vcov = vcovHC(model3, type = "HC0"))
ols_r4 <- coeftest(model4, vcov = vcovHC(model4, type = "HC0"))

stargazer(
  model1, model2, model3, model4,
  type             = "html",
  dep.var.caption  = "",
  out              = "Regression table.html",
  se               = list(ols_r1[, "Std. Error"], ols_r2[, "Std. Error"],
                          ols_r3[, "Std. Error"], ols_r4[, "Std. Error"]),
  dep.var.labels   = c("Inflation", "Core Inflation"),
  covariate.labels = c("Diesel", "Oil Prices (WTI)", "Spread Diesel-WTI", "Covid")
)
```
